Introduction to Markov state modeling with the PyEMMA software --- v\versionnumber

### 2018-11-14 manuscript/manuscript.tex
[manuscript] simplify HMM section, restructure citations

An alternative, which is much less sensitive to poor discretization,
is to estimate a hidden Markov model (HMM)~\cite{hmm-baum-welch-alg,hmm-tutorial,jhp-spectral-rate-theory,noe-proj-hid-msm,bhmm-preprint}.
HMMs are less sensitive to the discretization error as they sidestep the assumption of Markovian dynamics in the discretized space (illustrated in Fig.~\ref{fig:hmm-scheme}).
Instead, HMMs assume that there is an underlying (hidden) dynamic process which is Markovian
and gives rise to our observed data, e.g., the ($n$~states) discretized trajectories $s(t)$.
This is a powerful principle as we know that there is indeed an underlying process which is Markovian:
our molecular dynamics trajectories.

To estimate an HMM, we need a spectral gap after the $m^\textrm{th}$ timescale;
in practice, a timescale separation of $t_m \geq 2t_{m+1}$ is sufficient~\cite{pyemma}.
The HMM then consists of a transition matrix $\tilde{\mathbf{P}}(\tau)$ between $m<n$ hidden states
and a row-stochastic matrix ($\bm{\chi}$) of probabilities $\chi\left( s \middle| \tilde{s} \right)$
to emit the discrete state $s$ conditional on being in the hidden state $\tilde{s}$.

We can further compute a reversal of the emission matrix $\bm{\chi}\in\mathbb{R}^{m \times n}$:
the membership matrix $\mathbf{M}\in\mathbb{R}^{n \times m}$ which encodes
a fuzzy assignment of each of the $n$ observable microstates $s$ to the $m$ hidden states $\tilde{s}$ and,
thus, defines the \emph{coarse graining} of microstate.

>>>>

An alternative, which is much less sensitive to poor discretization,
is to estimate a hidden Markov model (HMM)~\cite{hmm-baum-welch-alg,jhp-spectral-rate-theory,noe-proj-hid-msm,bhmm-preprint}.
HMMs are less sensitive to the discretization error as they sidestep the assumption of Markovian dynamics in the discretized space (illustrated in Fig.~\ref{fig:hmm-scheme}).
Instead, HMMs assume that there is an underlying (hidden) dynamic process which is Markovian
and gives rise to our observed data, e.g., the ($n$~states) discretized trajectories $s(t)$.
This is a powerful principle as we know that there is indeed an underlying process which is Markovian:
our molecular dynamics trajectories.

To estimate an HMM, we need a spectral gap after the $m^\textrm{th}$ timescale;
in practice, a timescale separation of $t_m \geq 2t_{m+1}$ is sufficient~\cite{pyemma}.
The HMM then consists of a transition matrix $\tilde{\mathbf{P}}(\tau)$ between $m<n$ hidden states
and a row-stochastic matrix ($\bm{\chi}$) of probabilities $\chi\left( s \middle| \tilde{s} \right)$
to emit the discrete state $s$ conditional on being in the hidden state $\tilde{s}$.
For more details about HMM properties and the estimation algorithm, we suggest Ref.~\cite{hmm-tutorial}.
